Quasi-Monte Carlo Methods: What, Why, and How?

### 2025-01-26 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

The meaning of the coordinate weights is that those coordinates with a small weight should not contribute much to the variation.  In \eqref{eq:wtctrvariation}, if the $\gamma_{\ell_1}, \ldots, \gamma_{\ell_s}$ are small, then $\prod_k \gamma_{\ell_k}^{-2}$ will be large.  For $f$ to have modest variation requires $\lVert D^{\bsell}f \rVert_2$ to be small.  Even though a function may have a large nominal dimension, $d$, its effective dimension, i.e., the number of variables that substantially contribute to he variation, must be relatively few, so that the variation under coordinate weights is modest.

@@

For sample means where the nodes, $\{\bsx_i\}_{i=0}^{n-1}$ are randomized, one can define a measure of their quality based on the root mean squared error of $\hat{\mu}_n(f)$:
\begin{align}
    \text{RMSE}(\hat{\mu}_n(f)) & : = \sqrt{\mathbb{E}[(\mu(f) - \hat{\mu}_n(f))^2]} \\
    &\le \underbrace{ \sup_{\text{variation}(f,K) \le 1} \sqrt{\mathbb{E}[(\mu(f) - \hat{\mu}_n(f))^2]}}_{\text{RERR}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)} \, \text{variation}(f,K).
\end{align}
This quality measure $\text{RERR}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ is typically more optimisti, i.e., smaller, than the root mean squared of  $\text{discrepancy}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ because the latter is finding a worst case function for each random instance of nodes, while the former is finding one worst function.

@@

When approximating the true $\mu$ with $\hat{\mu}_n$, stopping criterion determine how large $n$ should be so that the error $\lvert \mu - \hat{\mu}_n \rvert$ is below an error tolerance $\epsilon$ either certainly or with sufficiently low uncertainty $\alpha$. The error tolerance and probability threshold are problem dependent. For example, in option pricing one may set the error tolerance to a penny, $\varepsilon=0.01$, and wish to find error bounds which hold with 95\% confidence, $\alpha=0.05$.

For classic MC with IID points, the Central Limit Theorem (CLT) may be used to get an approximate confidence interval $(\mu^-,\mu^+)$ where $\mu^\pm = \hat{\mu} \pm Z^{1-\alpha/2} \sigma/\sqrt{n}$, $\sigma = \mathrm{Std}[f(X)]$ and $Z^{1-\alpha/2}$ is the $1-\alpha/2$ quantile of the standard Gaussian distribution. If 
\begin{equation}
    n > (2 Z^{1-\alpha/2} \sigma/\varepsilon)^2
    \label{eq:sc_clt_iid_n}
\end{equation} then $\mu \in (\mu^-,\mu^+)$ with probability $1-\alpha$. A simple IID stopping criterion uses an initial sample of $n_0$ points to approximate $\sigma^2 \approx (n_0-1)^{-1} \sum_{i=0}^{n_0-1}(f(\bsx_i)-\hat{\mu})^2$, and then plugs this into \eqref{eq:sc_clt_iid_n} to determine $n$. As the CLT holds only as $n$ goes to infinity, this is an approximate stopping criterion. For integrands $f$ with bounded kurtosis, a guaranteed version of this two-stage-sampling idea has been developed to account for the finite sample size using Berry-Esseen inequalities \cite{JiaHic16a}.

A straightforward qMC stopping criterion may also be derived from the CLT using a few randomizations of a LD point set. Let $\bsx_{ir}$ be the $i^\mathrm{th}$ point in the $r^\mathrm{th}$ randomization of a LD point set with $0 \leq i < n$ and $1 \leq r \leq R$. For example, we may use $R$ independent shifts of a lattice point set or independent digital shifts of a digital net. Crucially, the randomized point sets presented in Section \ref{sec:random} are also LD. Let $\hat{\mu}_r = n^{-1} \sum_{i=0}^{n-1} f(\bsx_{ir})$ be approximate means for each randomization and set $\hat{\mu} = R^{-1} \sum_{r=1}^R \hat{\mu}_r$ to be the aggregate approximation. In the spirit of CLT, a standard approximate $1-\alpha$ confidence interval $(\mu^-,\mu^+)$ has $\mu^\pm = \hat{\mu} \pm t_{R-1}^{1-\alpha/2} S/\sqrt{R}$ where $S^2 = (R-1)^{-1} \sum_{r=1}^R (\hat{\mu}_r - \hat{\mu})^2$ and $t_{R-1}^{1-\alpha/2}$ is the $1-\alpha/2$ quantile of the Student $t$ distribution with $R-1$ degrees of freedom. Here we have used the Student $t$ distribution since $R$ is typically small, for example $R=15$. To use this confidence interval for a stopping criterion, one may iteratively double the sample size $n$ until the interval width is sufficiently low. In \cite{l2023confidence}, comprehensive numerical experiments show that this straightforward qMC stopping criterion achieve better coverage than more involved bootstrap $t$ methods for a variety of test functions.

@@

Another idea which avoids multiple randomizations is to treat $f$ as a Gaussian process, as mentioned in Section \ref{sec:randBayes}. Conditioned on observations $f(\bsx_i)$, a closed form expression exists for the $1-\alpha$ credible intervals for $\mu$. While also applicable when using IID points, using LD sequences paired with special kernels enables these credible intervals to be computed at only $\calO(n \log n)$ cost \cite{rathinavel2019fast}. Again we may iteratively double the sample size $n$ until the credible intervals are sufficiently tight.

Specifically, pairing shifted lattices with shift-invariant kernels as defined in \eqref{eq:shinvK} yields circulant Gram matrices for which computations can be done quickly using the fast Fourier transform \cite{RatHic19a}. Similarly, pairing digitally shifted digital nets with digitally-shift-invariant kernels yields recursive-block Toeplitz Gram matrices for which computations can be done quickly using fast Walsh-Hadamard transform \cite{JagHic22a}.

@@

Often (quasi-)Monte Carlo problems may be written as $\mu = \bbE[g(\bsT)]$ where $\bsT$ is a $d$-dimensional random variable with density $\lambda$. Since qMC points have LD with the standard uniform distribution $\bsX \sim \calU[0,1]^d$, we perform an invertible variable transformation $\bsPsi: [0,1]^d \to \calM$ satisfying $\mathrm{range}(\bsT) \subseteq \calM$. This implies
$\mu = \bbE[f(\bsX)]$ where  
$$f(\bsX) = g(\bsPsi(\bsX)) \frac{\lambda (\bsPsi(\bsX))}{\nu(\bsPsi(\bsX))}$$
and $\nu$ is the density of $\bsPsi(\bsX)$. Specifically, $\nu(\bsPsi(\bsX))= \lvert \bsJ_\bsPsi(\bsX) \rvert^{-1}$ where $\lvert \bsJ_\bsPsi(\bsX) \rvert$ is the determinant of the Jacobian of $\bsPsi$.

The above formulation is equivalent to importance sampling by a distribution with density $\nu$. Optimally, we would want to choose $\nu \propto g\lambda$ so that $f$ is nearly constant, but this is typically infeasible in practice due to our limited knowledge of $g$. Choosing $\bsPsi$ so that $\bsT = \bsPsi(\bsX)$ gives the simplification $f(\bsX) = g(\bsPsi(\bsX))$. An example of this was given in Section \ref{sec:practice} where $\bsT \sim \calN(\boldsymbol{0},\mathsf{I}/2)$ and $\bsPsi(\bsX) = \Phi^{-1}(\bsX)/\sqrt{2}$ where $\Phi^{-1}$ is the inverse CDF of the standard Gaussian distribution applied elementwise.

As observed in Section \ref{sec:coordwts}, qMC is more effective if the variation of $f$ is concentrated in the lower indexed coordinates.  For finance examples where the outcome depends more on the gross behavior of the underlying stochastic process, rather than the high frequency behavior, it makes sense to choose the transformation accordingly.  For example, if $\bsT$ represents a Brownian motion, then a principal component analysis coupled with the inverse Gaussian distribution shows greater convergence gains.

@@

Let $\{Y_l := f_l(\bsX_{l})\}_{l=1}^L$ be a sequence of approximations to $Y = f(\bsX)$, where $\bsX_l \sim \mathcal{U}[0,1)^{d_l}$.  Here we assume that $d_1 < \cdots < d_L$, and that the cost of evaluating $f_l(\bsx_{l})$ is $\mathcal{O}(d_l)$.  Also, $d_L = d$, $Y_{d_L} = Y$, and $f_L = f$. Then
\begin{align*}
    \mu & := \mathbb{E}(Y) = \mathbb{E}\bigl[f(\bsX)\bigr]
    = \mathbb{E}(Y_1) + \mathbb{E}(Y_2 - Y_1) + \cdots + \mathbb{E}(Y_{d_L} - Y_{d_{L-1}}) \\
    & = \int_{[0,1)^{d_1}} f_1(\bsx^{(1)}) \, \mathrm{d} \bsx^{(1)} +  \int_{[0,1)^{d_2}} \bigl[f_2(\bsx^{(2)}) - f_1(\bsx^{(1)})\bigr] \, \mathrm{d} \bsx^{(2)} + \cdots \\
    & \qquad \qquad + \int_{[0,1)^{d_L}} \bigl[f_L(\bsx^{(L)}) - f_{L-1}(\bsx^{(L-1)}) \bigr] \, \mathrm{d} \bsx^{(L)},
\end{align*}
where $\bsx^{(1)}  = (x_1, \dots, x_{d_1}), \ldots, \bsx^{(L)} = (x_1, \dots, x_{d_L})$.
This series of integrals may be approximated by a series of sample means, namely
\begin{multline} \label{eq:MLmuhat}
    \hat{\mu}^{\text{ML}}_{\bsn,\bsd}  := \frac{1}{n_1}  \sum_{i=0}^{n_1-1} f_1(\bsx^{(1)}_i)  +  \frac{1}{n_2}  \sum_{i=0}^{n_2-1}  \bigl[f_2(\bsx^{(2)}_i) - f_1(\bsx^{(1)}_i)\bigr] + \cdots \\
     + \frac{1}{n_L}  \sum_{i=0}^{n_L-1}  \bigl[f_L(\bsx_i^{(L)}) - f_{L-1}(\bsx_i^{(L-1)}) \bigr].
\end{multline}
If $C_1$ is the computational cost of $f^{(1)}(\bsx^{(1)})$ and $C_l$ is the computational cost of $f^{(l)}(\bsx^{(1)})$ of $f_L(\bsx_i^{(l)}) - f_{l-1}(\bsx_i^{(l-1)})$ for $l = 2, \ldots, L$, then the total computational cost of 
\eqref{eq:MLmuhat} is 
\begin{equation} \label{eq:MLcostdef}
    \text{cost}(\bsn,\bsd) = n_1 C_1 + \cdots + n_L C_L.
\end{equation}

@@

If the $V_l$ decay quickly enough, which means that the corrections $f_l - f_{l-1}$ depending on many variables decay quickly enough, in particular, if $r > d_1$, then 
\begin{equation}
         \text{cost}(\bsn,\bsd) \le \frac{\alpha \beta d_1} {r \varepsilon \bigl( 1 - (d_1/r)^{1/2}\bigr)^2} = \mathcal{O}(d_1/\varepsilon).
\end{equation}
In this case $\text{cost}(\bsn,\bsd)$ for this high dimensional problem is ont much worse than the cost of ensuring that just error of the rough $d_1$ approximation is smaller than $\varepsilon$, which would be $V_1C_1/\varepsilon \le \alpha \beta d_1/(r\varepsilon)$.

@@

%\section{Ongoing Research} \label{sec:ongoing}
\section{Conclusion} \label{sec:conclusion}
Quasi-Monte Carlo methods, based on low discrepancy sequences, can approximate to the mean of a function, $f$, of $d$ variables to within an error tolerance of $\varepsilon$ at a cost of roughly $\mathcal{O}(\varepsilon^{-1})$ function evaluations, provided that the problem is formulated so that $f$ depends primarily on a modest number of variables.  This means that qMC outperforms alternative methods such as simple Monte Carlo and sampling on grids.  Using the multilevel qMC, the cost to reach the error tolerance is often essentially the same for $d=1$ and $d$ unbounded.  Software libraries are available to facilitate qMC computations.

>>>>

The meaning of the coordinate weights is that those coordinates with a small weight should not contribute much to the variation.  In \eqref{eq:wtctrvariation}, if the $\gamma_{\ell_1}, \ldots, \gamma_{\ell_s}$ are small, then $\prod_k \gamma_{\ell_k}^{-2}$ will be large.  For $f$ to have modest variation requires $\lVert D^{\bsell}f \rVert_2$ to be small.  Even though a function may have a large nominal dimension, $d$, its effective dimension, i.e., the number of variables that substantially contribute to the variation, must be relatively few, so that the variation under coordinate weights is modest.

@@

For sample means where the nodes, $\{\bsx_i\}_{i=0}^{n-1}$ are randomized, one can define a measure of their quality based on the root mean squared error of $\hat{\mu}_n(f)$:
\begin{align}
    \text{RMSE}(\hat{\mu}_n(f)) & : = \sqrt{\mathbb{E}[(\mu(f) - \hat{\mu}_n(f))^2]} \\
    &\le \underbrace{ \sup_{\text{variation}(f,K) \le 1} \sqrt{\mathbb{E}[(\mu(f) - \hat{\mu}_n(f))^2]}}_{\text{RERR}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)} \, \text{variation}(f,K).
\end{align}
This quality measure $\text{RERR}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ is typically more optimistic, i.e., smaller, than the root mean squared of  $\text{discrepancy}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ because the latter is finding a worst case function for each random instance of nodes, while the former is finding one worst function.

@@

When approximating the true $\mu$ with $\hat{\mu}_n$, stopping criterion determine how large $n$ should be so that the error $\lvert \mu - \hat{\mu}_n \rvert$ is below an error tolerance $\epsilon$ either certainly or with sufficiently low uncertainty $\alpha$. The error tolerance and probability threshold are problem dependent. For example, in option pricing one may set the error tolerance to a penny, $\varepsilon=0.01$, and wish to find error bounds which hold with 95\% confidence, $\alpha=0.05$. The stopping criteria presented here are implemented in QMCPy \cite{QMCPy2020a}.

For simple Monte Carlo with IID points, the Central Limit Theorem (CLT) may be used to get an approximate confidence interval $(\mu^-,\mu^+)$ where $\mu^\pm = \hat{\mu} \pm Z^{1-\alpha/2} \sigma/\sqrt{n}$, $\sigma = \mathrm{Std}[f(X)]$ and $Z^{1-\alpha/2}$ is the $1-\alpha/2$ quantile of the standard Gaussian distribution. If 
\begin{equation}
    n > (2 Z^{1-\alpha/2} \sigma/\varepsilon)^2
    \label{eq:sc_clt_iid_n}
\end{equation} then $\mu \in (\mu^-,\mu^+)$ with probability $1-\alpha$. A simple IID stopping criterion uses an initial sample of $n_0$ points to approximate $\sigma^2 \approx (n_0-1)^{-1} \sum_{i=0}^{n_0-1}(f(\bsx_i)-\hat{\mu})^2$, and then plugs this into \eqref{eq:sc_clt_iid_n} to determine $n$. As the CLT holds only as $n$ goes to infinity, this is an approximate stopping criterion. For integrands $f$ with bounded kurtosis, a guaranteed version of this two-stage-sampling idea has been developed to account for the finite sample size using  Berry-Esseen inequalities\cite{JiaHic16a}.

A straightforward qMC stopping criterion may also be derived from the CLT using a few randomizations of a LD point set. Let $\bsx_{ir}$ be the $i^\mathrm{th}$ point in the $r^\mathrm{th}$ randomization of a LD point set with $0 \leq i < n$ and $1 \leq r \leq R$. For example, we may use $R$ independent shifts of a lattice point set or independent digital shifts of a digital net. Crucially, the randomized point sets presented in Section \ref{sec:random} are also LD. Let $\hat{\mu}_r = n^{-1} \sum_{i=0}^{n-1} f(\bsx_{ir})$ be approximate means for each randomization and set $\hat{\mu} = R^{-1} \sum_{r=1}^R \hat{\mu}_r$ to be the aggregate approximation. In the spirit of CLT, a standard approximate $1-\alpha$ confidence interval $(\mu^-,\mu^+)$ has $\mu^\pm = \hat{\mu} \pm t_{R-1}^{1-\alpha/2} S/\sqrt{R}$ where $S^2 = (R-1)^{-1} \sum_{r=1}^R (\hat{\mu}_r - \hat{\mu})^2$ and $t_{R-1}^{1-\alpha/2}$ is the $1-\alpha/2$ quantile of the Student $t$ distribution with $R-1$ degrees of freedom. Here we have used the Student $t$ distribution since $R$ is typically small, for example $R=15$. To use this confidence interval for a stopping criterion, one may iteratively double the sample size $n$ until the interval width is sufficiently low. In \cite{l2023confidence}, comprehensive numerical experiments show that this straightforward qMC stopping criterion achieve better coverage than more involved bootstrap $t$ methods across a variety of test functions.

@@

Another idea which avoids multiple randomizations is to treat $f$ as a Gaussian process, as mentioned in Section \ref{sec:randBayes}. Conditioned on observations $f(\bsx_i)$, a closed form expression exists for the $1-\alpha$ credible intervals for $\mu$. While also applicable when using IID points, using LD sequences paired with special kernels enables these credible intervals to be computed at only $\calO(n \log n)$ cost \cite{rathinavel2019fast}. Again one may iteratively double the sample size $n$ until the credible intervals are sufficiently tight.

Specifically, pairing shifted lattices with shift-invariant kernels as defined in \eqref{eq:shinvK} yields circulant Gram matrices for which computations can be done quickly using the fast Fourier transform \cite{RatHic19a}. Similarly, pairing digitally shifted digital nets with digitally-shift-invariant kernels yields recursive-block Toeplitz Gram matrices \AGSNote{Different name?} for which computations can be done quickly using the fast Walsh-Hadamard transform \cite{JagHic22a}.

@@

Often qMC problems may be written as $\mu = \bbE[g(\bsT)]$ where $\bsT$ is a $d$-dimensional random variable with density $\lambda$. Since qMC points have LD with the standard uniform distribution $\bsX \sim \calU[0,1]^d$, we perform an invertible variable transformation $\bsPsi: [0,1]^d \to \calM$ satisfying $\mathrm{range}(\bsT) \subseteq \calM$. This implies
$\mu = \bbE[f(\bsX)]$ where  
$$f(\bsX) = g(\bsPsi(\bsX)) \frac{\lambda (\bsPsi(\bsX))}{\nu(\bsPsi(\bsX))}$$
and $\nu$ is the density of $\bsPsi(\bsX)$. Specifically, $\nu(\bsPsi(\bsX))= \lvert \bsJ_\bsPsi(\bsX) \rvert^{-1}$ where $\lvert \bsJ_\bsPsi(\bsX) \rvert$ is the determinant of the Jacobian of $\bsPsi$.

The above formulation is equivalent to importance sampling \AGSNote{reference} by a distribution with density $\nu$. Optimally, one would choose $\nu \propto g\lambda$ so that $f$ is constant, but this is typically infeasible in practice due to our limited knowledge of $g$. Choosing $\bsPsi$ so that $\bsT = \bsPsi(\bsX)$ gives the simplification $f(\bsX) = g(\bsPsi(\bsX))$. An example of this was given in Section \ref{sec:practice} where $\bsT \sim \calN(\boldsymbol{0},\mathsf{I}/2)$ and $\bsPsi(\bsX) = \Phi^{-1}(\bsX)/\sqrt{2}$ where $\Phi^{-1}$ is the inverse CDF of the standard Gaussian distribution applied elementwise.

As observed in Section \ref{sec:coordwts}, qMC is more effective if the variation of $f$ is concentrated in the lower indexed coordinates.  For finance examples where the outcome depends more on the gross behavior of the underlying stochastic process, rather than the high frequency behavior, it makes sense to choose the transformation accordingly.  For example, if $\bsT$ represents a Brownian motion, then a principal component analysis coupled with the inverse Gaussian distribution shows greater convergence gains \AGSNote{compared to the Cholesky decomposition or bridge construction?}.

@@

Let $\{Y_l := f_l(\bsX_{l})\}_{l=1}^L$ be a sequence of approximations to $Y = f(\bsX)$, where $\bsX_l \sim \mathcal{U}[0,1)^{d_l}$.  Here we assume that $d_1 < \cdots < d_L$, and that the cost of evaluating $f_l(\bsx_{l})$ is $\mathcal{O}(d_l)$.  Also, $d_L = d$, $Y_{d_L} = Y$, and $f_L = f$. Then
\begin{align*}
    \mu & := \mathbb{E}(Y) = \mathbb{E}\bigl[f(\bsX)\bigr]
    = \mathbb{E}(Y_1) + \mathbb{E}(Y_2 - Y_1) + \cdots + \mathbb{E}(Y_{d_L} - Y_{d_{L-1}}) \\
    & = \int_{[0,1)^{d_1}} f_1(\bsx^{(1)}) \, \mathrm{d} \bsx^{(1)} +  \int_{[0,1)^{d_2}} \bigl[f_2(\bsx^{(2)}) - f_1(\bsx^{(1)})\bigr] \, \mathrm{d} \bsx^{(2)} + \cdots \\
    & \qquad \qquad + \int_{[0,1)^{d_L}} \bigl[f_L(\bsx^{(L)}) - f_{L-1}(\bsx^{(L-1)}) \bigr] \, \mathrm{d} \bsx^{(L)},
\end{align*}
where $\bsx^{(1)}  = (x_1, \dots, x_{d_1}), \ldots, \bsx^{(L)} = (x_1, \dots, x_{d_L})$.
This series of integrals may be approximated by a series of sample means, namely
\begin{multline} \label{eq:MLmuhat}
    \hat{\mu}^{\text{ML}}_{\bsn,\bsd}  := \frac{1}{n_1}  \sum_{i=0}^{n_1-1} f_1(\bsx^{(1)}_i)  +  \frac{1}{n_2}  \sum_{i=0}^{n_2-1}  \bigl[f_2(\bsx^{(2)}_i) - f_1(\bsx^{(1)}_i)\bigr] + \cdots \\
     + \frac{1}{n_L}  \sum_{i=0}^{n_L-1}  \bigl[f_L(\bsx_i^{(L)}) - f_{L-1}(\bsx_i^{(L-1)}) \bigr].
\end{multline}
If $C_1$ is the computational cost of $f^{(1)}(\bsx^{(1)})$ and $C_l$ is the computational cost of \AGSNote{$f^{(l)}(\bsx^{(1)})$ of $f_L(\bsx_i^{(l)}) - f_{l-1}(\bsx_i^{(l-1)})$} for $l = 2, \ldots, L$, then the total computational cost of 
\eqref{eq:MLmuhat} is 
\begin{equation} \label{eq:MLcostdef}
    \text{cost}(\bsn,\bsd) = n_1 C_1 + \cdots + n_L C_L.
\end{equation}

@@

If the $V_l$ decay quickly enough, which means that the corrections $f_l - f_{l-1}$ depending on many variables decay quickly enough, in particular, if $r > d_1$, then 
\begin{equation}
         \text{cost}(\bsn,\bsd) \le \frac{\alpha \beta d_1} {r \varepsilon \bigl( 1 - (d_1/r)^{1/2}\bigr)^2} = \mathcal{O}(d_1/\varepsilon).
\end{equation}
In this case $\text{cost}(\bsn,\bsd)$ for this high dimensional problem is not much worse than the cost of ensuring that just error of the rough $d_1$ approximation is smaller than $\varepsilon$, which would be $V_1C_1/\varepsilon \le \alpha \beta d_1/(r\varepsilon)$.

@@

%\section{Ongoing Research} \label{sec:ongoing}
\section{Conclusion} \label{sec:conclusion}
Quasi-Monte Carlo methods, based on low discrepancy sequences, can approximate the mean of a function, $f$, of $d$ variables to within an error tolerance of $\varepsilon$ at a cost of roughly $\mathcal{O}(\varepsilon^{-1})$ function evaluations, provided that the problem is formulated so that $f$ depends primarily on a modest number of variables.  This means that qMC outperforms alternative methods such as simple Monte Carlo and sampling on grids.  Using the multilevel qMC, the cost to reach the error tolerance is often essentially the same for $d=1$ and $d$ unbounded.  Software libraries are available to facilitate qMC computations.
